---
title: "Homework 3"
Author: Ayeshra Acharya 
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
```

##Problem 1 
```{r}
data('instacart')
```

The `instacart` dataset contains `r nrow(instacart)` observations (representing products from orders) and `r ncol(instacart)` variables. The key variables in the `instacart` dataset include `reordered`, which describes if this product has been ordered by this user in the past,`order_dow`, which describes the day in which the order was placed, and `order_hour_of_day`, the hour of the day on which the order was placed. There are also item variables - name, aisle, department, and some numeric codes. 

How many aisles and which are most items from?
```{r}
instacart %>%
  count(aisle)%>%
  arrange(desc(n))
```
The aisles that the most items are ordered from are "fresh vegetables" (150609 orders) and "fresh fruit" (150473 orders).

Making a plot 
```{r}
instacart %>%
  count(aisle) %>%
  filter (n>1000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,n)
  ) %>%
  ggplot(aes(x=aisle,y=n)) +
  geom_point()
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1 ))
```
Making a table 
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle,rank) %>%
  knitr::kable()
```

## Problem 2
Loading, tidying and wrangling data: 
```{r}
accel_data = 
  read_csv("./data/accel_data.csv",col_names = TRUE ) %>%
  janitor::clean_names() %>%
  mutate(
    weekend = recode(day, "Saturday" = 1, "Sunday" = 1, "Monday" = 0, "Tuesday" = 0, "Wednesday" = 0, "Thursday" = 0, "Friday" = 0),
    weekday = recode(day, "Saturday" = 0, "Sunday" = 0, "Monday" = 1, "Tuesday" = 1, "Wednesday" = 1, "Thursday" = 1, "Friday" = 1),
    day = ordered(day, c("Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday")),
    week = ordered(week, c("1", "2", "3", "4", "5"))) %>% 
  select(week, day_id, day, weekend, weekday, everything()) %>% 
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = "activity_minute_num",
    names_prefix = "activity_",
    values_to = "activity_counts") %>% 
  mutate(activity_minute_num = as.numeric(activity_minute_num))
```
The `accel_data` dataset contains `r nrow(accel_data)` observations and `r ncol(accel_data)` variables. The `r ncol(accel_data)` variables in the dataset are `week`, `day_id`, `day`, `weekend`, `weekday`, `activity_minute_num`, and `activity_counts`.

## Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals, check for trends. 
```{r}
total_act = accel_data %>% 
  group_by(week, day_id, day) %>% 
  summarize (total_min = sum(activity_counts))

knitr::kable(total_act)

trends_wk = total_act %>% 
  ggplot(aes(x = day, y = total_min, color = week, group = week)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(~week) +
  labs(
    title = "Total daily activity over 5 weeks",
    x = "Day of week",
    y = "Total activity per day (min)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

trends_wk

trends_day = total_act %>% 
  ggplot(aes(x = week, y = total_min, color = day, group = day)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(~day) +
  labs(
    title = "Total daily activity over 5 weeks",
    x = "Week",
    y = "Total activity per day (min)") 
trends_day
```
There are no obvious trends or patterns that stand out when looking at the `accel_data` dataset. Once we take a closer look, we can see that for the first two weeks(week 1 and week 2), the total activity per day is lowest on Mondays and then increases throughout the week. Additionally, for weeks 4 and 5, we can see that Saturdays and Sundays have the lowest total activity per day. 

##Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_plot = accel_data %>% 
  group_by(week, day_id, day) %>% 
  summarize (average_min = mean(activity_counts)) %>% 
  ggplot(aes(x = day_id, y = average_min, color = day)) +
  geom_line() +
  labs(
    title = "24-hour activity of Male Accelerometer Data Over 5-Week Period",
    x = "Day in 5-Week Period",
    y = "Average Activity in Minutes "
  )
accel_plot
```
Similar to the previous plot, we can see here that the average activity for Monday was very low, and increased throughout the week. However, from this plot we also notice that the average activity for Saturday and Sunday were high and then decreased throughout the week. 

##Problem 3 
```{r settings, message = FALSE}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)
library(hexbin)
```


```{r}
noa_data = 
  read_csv("./data/nynoaadat.csv", col_names = TRUE) %>%
  janitor::clean_names() %>%
  separate(date,c("year","month","day"),sep = "-") %>% 
  mutate(tmax = as.double(tmax) / 10, 
         tmin = as.double(tmin) / 10, 
         prcp = prcp / 10,
         month = month.name[as.integer(month)], 
         year = as.integer(year))
```
The `noa_data` dataset that we are using comes from The National Oceanic and Atmospheric Association(NOAA) and it provides us with summary statistics from weather stations in New York state. The `noa_data` dataset contains 2,595,176 rows (observations) and 7 columns (variables). The key variables in this dataset are: date, tmax, tmin, prcp and snow, indicating date, maximum temperature, precipitation, and snowfall. The variables prcp, snow, snwd, tmax, and tmin have missing data. Specifically, prcp is missing 145838 values, snow is missing 381221 values, snwd is missing 591786 values, tmax is missing 1134358 values, and tmin is missing 1134420 values. Missing data does seem to be a significant issue in this dataset because of the large number of values missing. 

For snowfall, what are the most commonly observed values? Why?
```{r}
noa_data %>% 
  group_by(snow) %>% 
  summarize(number_time_snow = n()) %>% 
  arrange(min_rank(desc(number_time_snow)))
noa_data %>% 
  group_by(snow, month) %>% 
  summarize(number_time_snow = n()) %>% 
  arrange(min_rank(desc(number_time_snow)))
```
For snowfall, the most commonly observed values are 0 and 25. This is interesting because most of the days out of the year, there is not snowfall - but, some days in the winter months experience heavy snow. 

## Two panel plot showing the average max temperature in January and in July in each station across years. 
```{r}
noa_data %>% 
  filter(month %in% c("January", "July")) %>% 
  group_by(year, month) %>% 
  summarize(avg_temp = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = avg_temp)) + 
    geom_line(color = "#7496D2", size = 1) + 
  facet_grid(~ month, 
             scales = "free") + 
  labs(title = "Figure 4: Average maximum temperature (°C) in January and July, 1981-2010", 
       x = "Year", 
       y = "Average Daily Temperature (°C)") +
  theme(axis.text.x = element_text(angle = 45,
                                   vjust = 0.4), 
        strip.background = element_rect(fill = "pink"), 
        strip.text = element_text(color = "black", 
                                  face = "bold"))
```
The temperatures for the month of January have more variability than for the month of July. There are not any outliers, but from the years 1985-1990, the month of July had low average maximum temperature.

##Make a two-panel plot showing (i) tmax vs tmin for the full dataset and  make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}
library(patchwork)
library(hexbin)
plot_tmax_tmin =
  noa_data %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()+
  labs(title = "tmax vs. tmin", x = "tmin (ºC)", y = "tmax (ºC)") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

plot_snow_year = 
  noa_data %>% 
  filter(0 < snow & snow < 100) %>%
  ggplot(aes( x = year, y = snow)) +
  geom_hex() + 
  scale_x_discrete(breaks = c("1981","1990","2000","2010")) +
  labs(title = "Yearly Snowfall", y = "snow fall (mm)")+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
plot_tmax_tmin + plot_snow_year
```


```



