---
title: "Homework 3"
Author: Ayeshra Acharya 
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
```

##Problem 1 
```{r}
data('instacart')
```

The `instacart` dataset contains `r nrow(instacart)` observations (representing products from orders) and `r ncol(instacart)` variables. The key variables in the `instacart` dataset include `reordered`, which describes if this product has been ordered by this user in the past,`order_dow`, which describes the day in which the order was placed, and `order_hour_of_day`, the hour of the day on which the order was placed. There are also item variables - name, aisle, department, and some numeric codes. 

How many aisles and which are most items from?
```{r}
instacart %>%
  count(aisle)%>%
  arrange(desc(n))
```
The aisles that the most items are ordered from are "fresh vegetables" (150609 orders) and "fresh fruit" (150473 orders).

Making a plot 
```{r}
instacart %>%
  count(aisle) %>%
  filter (n>1000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,n)
  ) %>%
  ggplot(aes(x=aisle,y=n)) +
  geom_point()
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1 ))
```
Making a table 
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle,rank) %>%
  knitr::kable()
```

## Problem 2
Loading, tidying and wrangling data: 
```{r}
accel_data = 
  read_csv("./data/accel_data.csv",col_names = TRUE ) %>%
  janitor::clean_names() %>%
  mutate(
    weekend = recode(day, "Saturday" = 1, "Sunday" = 1, "Monday" = 0, "Tuesday" = 0, "Wednesday" = 0, "Thursday" = 0, "Friday" = 0),
    weekday = recode(day, "Saturday" = 0, "Sunday" = 0, "Monday" = 1, "Tuesday" = 1, "Wednesday" = 1, "Thursday" = 1, "Friday" = 1),
    day = ordered(day, c("Saturday", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday")),
    week = ordered(week, c("1", "2", "3", "4", "5"))) %>% 
  select(week, day_id, day, weekend, weekday, everything()) %>% 
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = "activity_minute_num",
    names_prefix = "activity_",
    values_to = "activity_counts") %>% 
  mutate(activity_minute_num = as.numeric(activity_minute_num))
```
The `accel_data` dataset contains `r nrow(accel_data)` observations and `r ncol(accel_data)` variables. The `r ncol(accel_data)` variables in the dataset are `week`, `day_id`, `day`, `weekend`, `weekday`, `activity_minute_num`, and `activity_counts`.

## Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals, check for trends. 
```{r}
total_act = accel_data %>% 
  group_by(week, day_id, day) %>% 
  summarize (total_min = sum(activity_counts))

knitr::kable(total_act)

trends_wk = total_act %>% 
  ggplot(aes(x = day, y = total_min, color = week, group = week)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(~week) +
  labs(
    title = "Total daily activity over 5 weeks",
    x = "Day of week",
    y = "Total activity per day (min)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

trends_wk

trends_day = total_act %>% 
  ggplot(aes(x = week, y = total_min, color = day, group = day)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(~day) +
  labs(
    title = "Total daily activity over 5 weeks",
    x = "Week",
    y = "Total activity per day (min)") 
trends_day
```
There are no obvious trends or patterns that stand out when looking at the `accel_data` dataset. Once we take a closer look, we can see that for the first two weeks(week 1 and week 2), the total activity per day is lowest on Mondays and then increases throughout the week. Additionally, for weeks 4 and 5, we can see that Saturdays and Sundays have the lowest total activity per day. 

##Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_plot = accel_data %>% 
  group_by(week, day_id, day) %>% 
  summarize (average_min = mean(activity_counts)) %>% 
  ggplot(aes(x = day_id, y = average_min, color = day)) +
  geom_line() +
  labs(
    title = "24-hour activity of Male Accelerometer Data Over 5-Week Period",
    x = "Day in 5-Week Period",
    y = "Average Activity in Minutes "
  )
accel_plot
```
Similar to the previous plot, we can see here that the average activity for Monday was very low, and increased throughout the week. However, from this plot we also notice that the average activity for Saturday and Sunday were high and then decreased throughout the week. 




